---
title: "R Notebook"
output: 
  html_notebook: 
    toc: TRUE
---

# R is so much more than a statistical program  
- RStudio as a file manager/git user interface  
- A graphic design tool  
- A webpage generator  
- A notebook/word processor
- A report generator
- The R console - sometimes this is all you have

# RStudio
- Sometimes RStudio is available on servers
- 4 panes  
- console vs. source  
- so many options for other panes  
- using RStudio in combination with projects creates a simple file structure for working with repositories in github, especially working directory.

# How to import data into R
## As a package
Some packages have datasets included, so you only have to "call" the data and it will appear
```{r}
library(ggplot2)

mtcars
```
## As a delimited file (csv or tsv) 
```{r}
library(readr) # included in library(tidyverse)
library(here)

file <- here("data", "Maalaea_Deep.csv")
read_csv(file)

# can also read from the web
read_csv("https://github.com/NCEAS/crescynt-training/blob/master/data/Maalaea_Deep.csv")
```

## As an excel file  
- You have to know which sheet of the file you want to bring in  
- You can't read straight from the web (...yet (v1.3.1), try after an update is released.)
```{r}
library(readxl)

file <- here("data", "2012_GPS-surveys.xlsx")

excel_sheets(file)

read_xlsx(path = file, sheet = "Data")

readxl::read_excel("https://github.com/pinskylab/leyteBuildDB/blob/master/data_raw/survey-data/2012_GPS-surveys.xlsx")

```

## As an RData file
```{r}
# can't from the web yet
readRDS("https://github.com/pinskylab/OceanAdapt/raw/master/data_clean/by_national.RData")

# can from a local file
file <- here("data", "DBdata.RData")

readRDS(file)
load(file)
```


